---
title: "GeoSpatial Chicago Crime Analysis 2001-Present"
author: "Ronald Washington III"
date: "March 13, 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing Libraries

```{r}
library(dplyr)
library(readr)
library(ggplot2) 
library(ggrepel)
library(knitr)
library(chron)
library(scales)
# library(sf)
# library(sp)
# library(raster)
# library(rgdal)
# library(RCurl)
# library(shapefiles)
# library(maptools)
```


```{r}
crime.data <- read.csv("https://data.cityofchicago.org/api/views/ijzp-q8t2/rows.csv?accessType=DOWNLOAD",nrows = 50000)
```

# Cleaning of data

```{r}
cleaned.crime.data <- subset(crime.data, !duplicated(crime.data$Case.Number)) #Removal of duplicates
cleaned.crime.data <- cleaned.crime.data %>% filter(Arrest=="true") #Limiting to only arrests 
cleaned.crime.data$Date <- as.POSIXct(cleaned.crime.data$Date,format=" %m/%d/%Y %H:%M") #readable date format
cleaned.crime.data$Time <- times(format(cleaned.crime.data$Date," %H:%M:%S")) #inclusion of time column

# Separating into time chucks for time analysis which will aid in seasonal analysis
Time.tag <- chron(times= c("00:00:00", "06:00:00", "12:00:00", "18:00:00","23:59:00"))
cleaned.crime.data$Time.tag <- cut(cleaned.crime.data$Time, breaks=Time.tag,labels=c("00-06","06-12", "12-18", "18-00"), include.lowest=TRUE)
cleaned.crime.data$Date <- as.POSIXct(strptime(cleaned.crime.data$Date,format= " %Y-%m-%d"))#readable date format / without time

# Removal of NAs found within Longitude and Latitutde
cleaned.crime.data <- subset(cleaned.crime.data, !is.na(cleaned.crime.data$Latitude))
```

## Reducing the number of crime types/descriptions
### Number of Different types of Crime type

```{r}
length(table(cleaned.crime.data$Primary.Type)) # 30 found

(table(cleaned.crime.data$Primary.Type)) #occurances of each
```

### Creating classifications for crime types

```{r}
cleaned.crime.data$crime <- as.character(cleaned.crime.data$Primary.Type)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("CRIM SEXUAL ASSAULT","PROSTITUTION", "SEX OFFENSE"), "SEXUAL", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("MOTOR VEHICLE THEFT"),"VEHICLE", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("GAMBLING", "INTERFERE WITH PUBLIC OFFICER","INTERFERENCE WITH PUBLIC OFFICER" ,"INTIMIDATION",
                                                     "LIQUOR LAW VIOLATION",  "OBSCENITY" , "NON-CRIMINAL","PUBLIC PEACE VIOLATION", "PUBLIC INDECENCY", 
                                                     "STALKING" ,  "NON-CRIMINAL (SUBJECT SPECIFIED)" ), "NON-VIOLATION", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime == "CRIMINAL DAMAGE", "DAMAGE",cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime == "CRIMINAL TRESPASS","TRESPASS", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("NARCOTICS","OTHER NARCOTIC VIOLATION","OTHER NARCOTIC VIOLATION"),
                                   "DRUG", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime ==  "DECEPTIVE PRACTICE","FRAUD", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("OTHER OFFENSE", "OTHER OFFENSE"), "OTHER", cleaned.crime.data$crime)

cleaned.crime.data$crime <- ifelse(cleaned.crime.data$crime %in% c("KIDNAPPING", "WEAPONS VIOLATION", "OFFENSE INVOLVING CHILDREN"), "VIOLATION", cleaned.crime.data$crime)
```

```{r}
summary(cleaned.crime.data)
```



## Aggregated Data Presenting Top 6 Types of Crimes and Relative Distribution 

```{r}
huh<- as.data.frame(cleaned.crime.data)
df_crime_daily <- huh %>%
  group_by(Date) %>%
  summarize(count = n()) %>%
  arrange(Date)
```

```{r}
df_category <- sort(table(huh$Primary.Type),decreasing = TRUE)
df_category <- data.frame(df_category[df_category > 1])
colnames(df_category) <- c("Category", "Frequency")
df_category$Percentage <- df_category$Frequency / sum(df_category$Frequency)
x<-head(df_category)
x
```



```{r}
df_theft <- huh %>% filter(grepl("ASSAULT", Primary.Type))

df_theft_daily <- df_theft %>%
  mutate(Date = as.Date(Date, "%m/%d/%Y")) %>%
  group_by(Date) %>%
  summarize(count = n()) %>%
  arrange(Date)


plot <- ggplot(df_theft_daily, aes(x = Date, y = count)) +
  geom_line(color = "#F2CA27", size = 0.1) +
  geom_smooth(color = "#1A1A1A") +
  # fte_theme() +
  scale_x_date(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  labs(x = "Date of Theft", y = "Number of NARCOTICS", title = "Daily NARCOTICS in San Francisco from ")
plot
```

### CREATE FUNCTION TO SHOW THE TRENDS OF THE TOP 6 CRIME TYPES

### CREATE HEATMAPS/CORRELATION MATRIXS SHOWING RELATIONSHIPS OF SEASONS

## Map showing crime in chicago

```{r}
library(leaflet)
data <- huh[1:10000,] # display the first 10,000 rows
data$popup <- paste("<b>Incident #: </b>", data$Case.Number, "<br>", "<b>Category: </b>", data$Primary.Type,
                    "<br>", "<b>Description: </b>", data$Description,
                    "<br>", "<b>Location Description: </b>", data$Location.Description,
                    "<br>", "<b>Community Area: </b>", data$Community.Area,
                    "<br>", "<b>Time: </b>", data$Time,
                    "<br>", "<b>Time: </b>", data$Year,
                    "<br>", "<b>Crime Type: </b>", data$crime,
                    "<br>", "<b>Longitude: </b>", data$Longitude,
                    "<br>", "<b>Latitude: </b>", data$Latitude)

leaflet(data, width = "100%") %>% addTiles() %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(provider = "Esri.WorldStreetMap",group = "World StreetMap") %>%
  addProviderTiles(provider = "Esri.WorldImagery",group = "World Imagery") %>%
  # addProviderTiles(provider = "NASAGIBS.ViirsEarthAtNight2012",group = "Nighttime Imagery") %>%
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = data$popup, clusterOptions = markerClusterOptions()) %>%
  addLayersControl(
    baseGroups = c("OSM (default)","World StreetMap", "World Imagery"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

