---
title: "new_crime_file"
author: "Ronald Washington III"
date: "March 19, 2019"
output: html_document
---


```{r setup, include =FALSE}

knitr::opts_chunk$set(echo = TRUE) 
```


```{r}
#  Comments by Ronald Washington III ----------------------------------------------------------------
# The code below was developing as a Senior Research project for DA 401 ("Seminar in Data Analytics")
# at Denison University. I was inspired to conduct research regarding my home-city Chicago, IL and 
# perform some investigatory analysis on the evolution of crime within the city from 2001-present day.
# The data utlized within my research was provided by the Chicago Police Department, and can be found
# at (https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2). This data includes
# information regarding important information such as date, type of crime, crime code, description, arrest 
# identification, police district, longitude and latitude, and etc. As I began my analysis for this project
# I gathered inspiration for my analysis from numerous other researcher that investigated crime in other 
# cities such as Baltimore, LA, New York, and San Francisco. (Each of these references will be listed below).
# From this investigation I intend to discover the distributions of various crimes occuring within Chicago, 
# and hopfully be able to map out each crime type. From there I wish to perform clustering and classification 
# methods to predict the types of crimes that can occur in the future. Furthermore, I am particularly interested
# in being able to recognize patterns of crime that occur in the city and potentially being able to identify 
# gang activtiy and its evolution in influencing Chicago Crime.
#
# STATE THE DATA SET USED TO IDENTIFY GANG ACTIVITY
#
# My primary objectives for the project were to (1) explore the data, (2) perform descriptive analysis and 
# various forms of clustering and classifying of crime, and (3) and ultimately visualization. 
#
# Ronald Washington III (GitHub: https://github.com/ron2297)

# File Descriptions -----------------------------------------------------------------------------------
#
# Crime Crime Data for 2001 through Present Day (2019) downloaded from Chicago Data Portal
# at (https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2) (n = 6823075)
```

# Importing Libraries
```{r include=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(ggforce)
library(ggplot2) 
library(ggrepel)
library(ggExtra)
library(ggmap)
library(ggthemes)
library(ggforce)
library(chron)
library(scales)
library(tidyr)
library(xts) 
library(lubridate)
library(highcharter)
library(treemap)
```




# Reading in Data
```{r eval=FALSE, include=FALSE}
#crime.data <- read.csv("C:/Users/Ron III/Desktop/Crimes_-_2001_to_present.csv") #,nrows = 500000)
crime.data <- read.csv(("C:/Users/Bloody Dachi/Documents/Crimes_-_2001_to_present (1).csv"))
#crime.data <- read.csv("https://data.cityofchicago.org/api/views/ijzp-q8t2/rows.csv?accessType=DOWNLOAD",nrows = 50000)
```



# Pre-processing - Cleaning of Crime Data

```{r eval=FALSE, include=FALSE}
df_Crime <- subset(crime.data, !duplicated(crime.data$Case.Number)) #Removal of duplicates
#Removing Unneeded Columns
df_Crime <- subset(df_Crime, select = -c(Case.Number,Description,Domestic,Beat,Ward,Community.Area,FBI.Code,X.Coordinate,Y.Coordinate))
#df_Crime <- df_Crime %>% filter(Arrest=="true") #Limiting to only arrests 
df_Crime$Date <- as.POSIXct(df_Crime$Date,format=" %m/%d/%Y %H:%M") #readable date format
df_Crime$Time <- times(format(df_Crime$Date," %H:%M:%S")) #inclusion of time column
df_Crime$DayOfWeek <-wday(as.Date(df_Crime$Date,'%m/%d/%Y'), label=TRUE)
df_Crime$Month <-months(as.Date(df_Crime$Date,'%m/%d/%Y'))

#as.POSIXct(df_Crime$Date,format=" %m/%d/%Y %H:%M") #readable date format
# Separating into time chucks for time analysis which will aid in seasonal analysis
Time.tag <- chron(times= c("00:00:00", "06:00:00", "12:00:00", "18:00:00","23:59:00"))
df_Crime$Time.tag <- cut(df_Crime$Time, breaks=Time.tag,labels=c("00-06","06-12", "12-18", "18-00"), include.lowest=TRUE)
df_Crime$Date <- as.POSIXct(strptime(df_Crime$Date,format= " %Y-%m-%d"))#readable date format / without time

# Removal of NAs found within Longitude and Latitutde and Date
df_Crime <- subset(df_Crime, !is.na(df_Crime$Latitude))
df_Crime <- subset(df_Crime, !is.na(df_Crime$Date))
```




## Reducing the number of crime types/descriptions
### Number of Different types of Crime type

```{r eval=FALSE, include=FALSE}
length(table(df_Crime$Primary.Type)) # 30 found

(table(df_Crime$Primary.Type)) #occurances of each
```


### Creating classifications for crime types

```{r eval=FALSE, include=FALSE}
df_Crime$crime <- as.character(df_Crime$Primary.Type)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("CRIM SEXUAL ASSAULT","PROSTITUTION", "SEX OFFENSE"), "SEXUAL", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("MOTOR VEHICLE THEFT"),"VEHICLE", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("GAMBLING", "INTERFERE WITH PUBLIC OFFICER","INTERFERENCE WITH PUBLIC OFFICER" ,"INTIMIDATION",
                                                     "LIQUOR LAW VIOLATION",  "OBSCENITY" , "NON-CRIMINAL","PUBLIC PEACE VIOLATION", "PUBLIC INDECENCY", 
                                                     "STALKING" ,  "NON-CRIMINAL (SUBJECT SPECIFIED)" ), "NON-VIOLATION", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime == "CRIMINAL DAMAGE", "DAMAGE",df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime == "CRIMINAL TRESPASS","TRESPASS", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("NARCOTICS","OTHER NARCOTIC VIOLATION","OTHER NARCOTIC VIOLATION"),
                                   "DRUG", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime ==  "DECEPTIVE PRACTICE","FRAUD", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("OTHER OFFENSE", "OTHER OFFENSE"), "OTHER", df_Crime$crime)

df_Crime$crime <- ifelse(df_Crime$crime %in% c("KIDNAPPING", "WEAPONS VIOLATION", "OFFENSE INVOLVING CHILDREN"), "VIOLATION", df_Crime$crime)
```


```{r}
#write.csv(df_Crime, file = "processs_crime_data.csv")
df_Crime <- read.csv(("C:/Users/Bloody Dachi/Documents/Chicago_Crime_Evolution_Analysis/Analysis_Folder/processs_crime_data.csv"))


```

### Removing Outliers from Longitude and Latitude

```{r}
df_Crime=df_Crime %>% filter(Latitude > 40) 
df_Crime=df_Crime %>% filter(Longitude > -90) 

#WantedData=df_Crime[df_Crime$Latitude %in% coors$df_Crime.Latitude, ]
#WantedData=df_Crime[df_Crime$Longitude %in% coors$df_Crime.Longitude, ]
summary(df_Crime$Longitude)
```

### Review of Simplified Crime Category Counts

```{r}
df_Crime <- subset(df_Crime, !is.na(df_Crime$crime))
(table(df_Crime$crime)) #occurances of each
```



# Data Exploration

## Aggregated Data Presenting Top 6 Types of Crimes and Relative Distribution 

```{r}
huh<- as.data.frame(df_Crime)
df_crime_daily <- huh %>%
  group_by(crime,Time.tag,DayOfWeek) %>%
  summarize(count = n()) %>%
  mutate(norm = count/sum(count)) %>%
  arrange(crime)
```



```{r}
df_category <- sort(table(huh$crime),decreasing = TRUE)
df_category <- data.frame(df_category[df_category > 1])
colnames(df_category) <- c("Category", "Frequency")
df_category$Percentage <- df_category$Frequency / sum(df_category$Frequency)
x<-head(df_category)
x #gathers top 6
```



# Visualization of Crimes and Arrest Relationship

```{r}
df_Crime$crime <- as.factor(df_Crime$crime)

by_Date <- df_Crime %>% group_by(Date) %>% summarise(Total = n())
tseries <- xts(by_Date$Total, order.by=as.POSIXct(by_Date$Date))

#Arrests_by_Date$Date[!(Arrests_by_Date$Date %in% by_Date$Date)]
## Creating timeseries of arrests made
Arrests_by_Date <- (df_Crime[df_Crime$Arrest == 'true',]) %>% group_by(Date) %>% summarise(Total = n())
arrests_tseries <- xts(Arrests_by_Date$Total, order.by=as.POSIXct(Arrests_by_Date$Date))


hchart(tseries, name = "crime") %>% 
 hc_add_series(arrests_tseries, name = "Arrest") %>%
 hc_add_theme(hc_theme_monokai()) %>%
 hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
 hc_title(text = "Trend of Chicago Crimes and Arrests") %>% 
 hc_legend(enabled = TRUE)
```



```{r fig.height=10, fig.width=12}
cat_value_freq <-  
  df_Crime %>% 
  select_if(is.factor) %>% 
  select_if(function(x) !is.ordered(x)) %>% 
  gather("var", "value") %>% 
  group_by(var) %>% 
  count(var, value) %>%
  mutate(prop = prop.table(n)) %>% 
  filter(prop > .02)

cat_plot1 <-
  ggplot(data = cat_value_freq,
       aes(x = reorder(stringr::str_wrap(value, 20), prop),
           y = prop)) +
  geom_bar(stat = "identity", fill = "tomato3") +
  coord_flip() +
  facet_wrap(~var,ncol=4, scales = "free") +
  ggthemes::theme_fivethirtyeight()

cat_plot1
```


```{r}
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
# %-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0
```





# Mapping Chicago

```{r}
chicago = get_stamenmap(bbox = c(left = -88.0225, bottom = 41.5949, right = -87.2713, top = 42.0677), maptype = c("toner-lite"), zoom = 13)
map = ggmap(chicago)
map
```



```{r}
# #map <- get_map("Chicago", zoom = 12, color = "bw")
# ggmap(map, extent='device') + 
#      geom_point(data=sample_n(df_Crime,100000), aes(x=X.Coordinate, y=Y.Coordinate),
#                 alpha = 1/10,color = "red")+
#      scale_colour_brewer(type="qual")
```




```{r}
chicago <- get_stamenmap(bbox = c(left = -88.0225, bottom = 41.5949, 
                                  right = -87.2713, top = 42.0677), 
                         zoom = 10)

ggmap(chicago)
```






```{r}
map_contours <- function(data_trunc, alp) {
        p1 = ggmap(chicago, extent='device') + 
        geom_point(data=data_trunc, aes(x=Longitude, y=Latitude), alpha= alp) + 
        stat_density2d(aes(x=Longitude, y=Latitude,
                           fill = ..level.., alpha = ..level..),
                size = 0.1, data = data_trunc, n=100,
                geom = "polygon") +
        theme(legend.position="none")
        return(p1)
}


plot_marginals <- function(data_trunc) {
        p2 = ggplot(data=data_trunc, aes(x=Longitude, y=Latitude), alpha=0.1)+
        geom_point()
        p2 = ggMarginal(p2 + theme_gray(), type = "histogram",
           fill = "steelblue", col = "darkblue")
        return(p2)
}
```





```{r}
p2 = plot_marginals(sample_n(df_Crime,10000))
plot(p2)
```



```{r}
p2 = map_contours(sample_n(df_Crime,10000),.1)
p2
```


# Crime vs Years
```{r}
data_plot = df_Crime %>%
        group_by(Year,Month) %>%
        summarise(count = n()) 
ggplot(data = data_plot,aes(group = Month, x=Year, y=count,col = Month)) + 
  geom_boxplot() + 
  ylab("Crime count in each month of a year")
```



  
```{r}
# top10_df <- df_Crime[df_Crime$crime %in% x[,1],]
# head(top10_df)
# top10_df
```




```{r}
ggmap(chicago, extent='device', legend="topleft") + 
  geom_point(data = df_Crime[df_Crime$crime %in% x[,1],], aes(x=Longitude, y=Latitude, col = crime), 
             size = 1.2, alpha = 0.5) + ggtitle("Mapping top 10 crimes")
```





```{r}
bbox = c(-88.0225,41.5949,-87.2713,42.0677)
chicago_map <- get_map(location = bbox, source = "stamen", maptype = "toner-lite")

#df_Crime$
# plot<-ggmap(chicago_map)+
#   geom_point(data = df_Crime,aes(x=Longitude,y=Latitude), color = "green", size = 0.5, alpha = 0.01)+theme_light()+labs(title = "ass")
# plot
```



#Top 10 is actually Top 6

```{r}
map_crime <- function(crime_df) {
  #filtered <- filter(crime_df, Primary.Type %in% crime)
  plot <- ggmap(chicago_map, extent='device') + 
          geom_density2d(data = crime_df%>% filter(Primary.Type %in% df_Crime[df_Crime$crime %in% x[,1],]$Primary.Type ) , 
          aes(x = Longitude, y = Latitude), size = 0.3) + stat_density2d(data = crime_df%>% filter(Primary.Type %in% df_Crime[df_Crime$crime %in% x[,1],]$Primary.Type ) , 
          aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01,
          bins = 16, geom = "polygon") + scale_fill_gradient(low = "green", high = "red", 
          guide = FALSE) + scale_alpha(range = c(0, 0.3), guide = FALSE)+
            facet_wrap_paginate(~ crime, ncol = 3,page =1)
  return(plot)
}
#length(unique(WantedData$crime))
# Display Top 6 Crime Types Desnity Distribution across Chicago
map_crime(df_Crime[1:1000000,])
```


