---
title: "Current_Statistics_Chicago_Crime"
author: "Ronald Washington III"
date: "April 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(dplyr) 
library(readr)
library(ggplot2) 
library(knitr)
library(plyr)
library(maptools)
library(rgdal)
library(rgeos)
library(shapefiles)
library(sf)
library(forecast)
library(sp)
library(spatstat)
```

# Reading in Data
```{r include=FALSE}
# Load Chicago Crimes 2016-2018 CSV
df_crime<-read.csv("C:/Users/Bloody Dachi/Desktop/processs_crime_data.csv")#,nrows = 50000)
#crime.data <- read.csv("https://data.cityofchicago.org/api/views/ijzp-q8t2/rows.csv?accessType=DOWNLOAD",nrows = 50000)
crimePPP_Assault$
```

## Filtering Data to only top 5 types of crimes 


```{r}               
df_crime_top5 <- subset(df_crime, crime %in% c("THEFT", "BATTERY", "DAMAGE" ,"DRUG","ASSAULT"))
```

# Cluster Identification Methods

[INSERT EXPLAINATION: WHY ARE YOU PERFORMING POINT PATTERN ANALYSIS - Due to your data being spatial you are interested in understanding the underlying patterns that occur from crimes in Chicago, and want to identifying the variablity between them and observe their temporal changes throughout time and attempt to predict off of it]

[INSERT EXPLAINATION OF WHAT POINT PATTERN ANALYSIS IS]

[INSERT EXPLAINATION OF HOW YOUR DATA IS INHOMOGENEOUS DUE TO THE INTENSITY (DENISTY) OF YOUR DATA CHANGING DEPENDING ON THE SPACE (LOCATION). STATE THAT YOU WILL BE USING INHOMOGENEOUS POISSON PROCESS]

[INSERT DESCRIPTION: YOU ARE INTERESTED IN ANALYZING THE TOP 5 TYPES OF CRIME IN CHICAGO ARE DISTRIBUTED IN SPACE AND TIME. SPECIFICALLY YOU WANT TO INVESTIGATE THE INTERACTION BETWEEN THESE TYPES OF CRIMES WITHIN POLICE DISTRICTS OR BETWEEN EACH OTHER]


Within this section we will be implementing numerous techniques to check for any form of clustering within our data points. Here we will be utilizing point pattern analysis as a metric of finding clustering, and we will be utilizing the Kolmogorov-Smirnov test as another method of determining if there are clusters.

## Point Pattern Analysis

### Removing Outlier Coordinates
```{r}
df_crime_top5=df_crime_top5 %>% filter(Latitude > 40) 
df_crime_top5=df_crime_top5 %>% filter(Longitude > -90) 
```

```{r include=FALSE}
coords <- SpatialPoints(df_crime_top5[,c("Longitude","Latitude")])
crime_spatial_df <-SpatialPointsDataFrame(coords,df_crime_top5)
proj4string(crime_spatial_df) <- CRS("+proj=longlat +ellps=WGS84")
```


```{r}
plot(crime_spatial_df)
```



```{r include=FALSE}

# Setting up Window Size / Formatting points to fit into shape of Chicago

#unzip("C:/Users/Ron III/Documents/GitHub/Chicago_Crime_Evolution_Analysis/Data_Sets_Files/Boundaries - ZIP Codes.zip", exdir = "C:/Users/Ron III/Documents/Crime_Research/shapefiles_1", overwrite = TRUE)

unzip("C:/Users/Bloody Dachi/Documents/Chicago_Crime_Evolution_Analysis/Data_Sets_Files/Boundaries - ZIP Codes.zip", exdir = "C:/Users/Bloody Dachi/Documents/Chicago_Crime_Evolution_Analysis/shapefiles_1", overwrite = TRUE)

chicago_shp <- readOGR(dsn = "C:/Users/Bloody Dachi/Documents/Chicago_Crime_Evolution_Analysis/shapefiles_1",layer = "geo_export_6c977322-0a01-4386-849c-278ec22209fe" )

w3=as.owin(chicago_shp)
```

### Checking for Inhomogeneity

[INSERT EXPLAINATION OF WHAT INHOMOGENEITY IS AND HOW IT IS REGARDING THE VARYING DENISTIES WITHIN THE QUADRAT PLOTS, IF IT WERE HOMOGENEOUS THEN ALL THE QUADRANTS WOULD BE THE SAME NUMBER]

#### Check #1 - Quadrat Distribution

```{r fig.height=8, fig.width=12}

combining_plots <- function(crime_type){
  crimePPP1 = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime==crime_type],crime_spatial_df$Latitude[crime_spatial_df$crime==crime_type],w3)
  par(mfrow=c(1,2))
  plot(crimePPP1, main = paste0("Point Plotting",crime_type), pch = ".", cols =2)
  qc <- quadratcount(crimePPP1) 
  plot(qc, main = paste0("Quadratcounting ",crime_type))
}
```

```{r fig.height=8, fig.width=12, warning=FALSE}
combining_plots("ASSAULT")
combining_plots("BATTERY")
combining_plots("DAMAGE")
combining_plots("DRUG")
combining_plots("THEFT")
```

[CONFIRM THAT ALL OF THE QUADRAT COUNTING IS INHOMOGENEOUS AND THAT WE CAN NOW MOVE FORWARD WITH ESTIMATING INTENSITY]

#### Creating Point Patterns for each Crime Type

```{r message=FALSE, warning=FALSE}
crimePPP_Theft = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime=="THEFT"],crime_spatial_df$Latitude[crime_spatial_df$crime=="THEFT"],w3)
crimePPP_Battery = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime=="BATTERY"],crime_spatial_df$Latitude[crime_spatial_df$crime=="BATTERY"],w3)
crimePPP_Damage = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime=="DAMAGE"],crime_spatial_df$Latitude[crime_spatial_df$crime=="DAMAGE"],w3)
crimePPP_Assault = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime=="ASSAULT"],crime_spatial_df$Latitude[crime_spatial_df$crime=="ASSAULT"],w3)
crimePPP_Drug = ppp(crime_spatial_df$Longitude[crime_spatial_df$crime=="DRUG"],crime_spatial_df$Latitude[crime_spatial_df$crime=="DRUG"],w3)

```

#### Labeling Crime Patterns 

```{r}
marks(crimePPP_Assault) = rep("Assualt", npoints(crimePPP_Assault))
marks(crimePPP_Theft) = rep("Theft", npoints(crimePPP_Theft))
marks(crimePPP_Battery) = rep("Battery", npoints(crimePPP_Battery))
marks(crimePPP_Damage) = rep("Damage", npoints(crimePPP_Damage))
marks(crimePPP_Drug) = rep("Drug", npoints(crimePPP_Drug))


```


#### Check #2 - Quadrat Monte Carlo Test / Visual Representation of Denisty Distribution

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
Inhomo_Test_Check <- function(crime_type){
  # the lambda argument specifys the density plot for assessing inhomogeneity
  print((Minhom<-quadrat.test(crime_type, nx = 4, ny = 10, method = "MonteCarlo", lambda = density(crime_type, sigma = .5))))
  plot(density(crime_type,sigma = .5), main =paste0("Grid counts for inhomogenious denisty: ",crime_type$marks[1]) )
  plot(Minhom, cex = 0.5, col = "white", add = T)
}

Inhomo_Test_Check(crimePPP_Theft)
Inhomo_Test_Check(crimePPP_Battery)
Inhomo_Test_Check(crimePPP_Damage)
Inhomo_Test_Check(crimePPP_Assault)
Inhomo_Test_Check(crimePPP_Drug)

```

[DISCUSS HOW THESE DIFFERENT DISTRUBITONS INSINUATES THAT THERE IS CLUSTERING IN THE POINT PATTERNS]



### Testing for Clustering within Crimes Patterns


#### Check #1 - Clark and Evans Test

```{r message=FALSE, warning=FALSE}
#The Clark and Evans (1954) aggregation index R 
#is a crude measure of clustering or ordering of 
#a point pattern. It is the ratio of the observed 
#mean nearest neighbour distance in the pattern to 
#that expected for a Poisson point process of the 
#same intensity. A value R > 1 suggests ordering, 
#while R < 1 suggests clustering.


Theft_Cluster_Check<- clarkevans(crimePPP_Theft, correction = "guard", clipregion = erosion.owin(rr,  r = .1))
Assualt_Cluster_Check<- clarkevans(crimePPP_Assault, correction = "guard", clipregion = erosion.owin(rr,  r = .1))
Battery_Cluster_Check<- clarkevans(crimePPP_Battery, correction = "guard", clipregion = erosion.owin(rr,  r = .1))
Drug_Cluster_Check<- clarkevans(crimePPP_Drug, correction = "guard", clipregion = erosion.owin(rr,  r = .1))
Damage_Cluster_Check<- clarkevans(crimePPP_Damage, correction = "guard", clipregion = erosion.owin(rr,  r = .1))

crime_cluster_check_df <- data.frame(Assualt = as.numeric(0),Theft = as.numeric(0),Battery = as.numeric(0),Drug = as.numeric(0),Damage = as.numeric(0))
  
crime_cluster_check_df$Theft <- Theft_Cluster_Check
crime_cluster_check_df$Assualt <- Assualt_Cluster_Check
crime_cluster_check_df$Battery <- Battery_Cluster_Check
crime_cluster_check_df$Drug <- Drug_Cluster_Check
crime_cluster_check_df$Damage <- Damage_Cluster_Check
crime_cluster_check_df
```

[DISCUSS HOW THE INDEX R VALUES FOR EACH OF THE POINT PATTERNS PRESENT THTA THERE IS CLUSTERING WITHIN THEIR DATA.]


#### Check #2 - Plotting G-Function to identify whether there is clustering in each crime pattern

```{r message=FALSE, warning=FALSE}
#The nearest neighbour distance distribution function (also called the ``event-to-event'' or ``inter-event'' distribution) of a point process X is the cumulative distribution function G the distance from a typical random point of X to the nearest other point of X.
Gest_Cluster_Graph <- function(crime_type){
  # the lambda argument specifys the density plot for assessing inhomogeneity
  G <- Gest(crime_type)
  plot(G)
}

Gest_Cluster_Graph(crimePPP_Theft)
Gest_Cluster_Graph(crimePPP_Battery)
Gest_Cluster_Graph(crimePPP_Damage)
Gest_Cluster_Graph(crimePPP_Assault)
Gest_Cluster_Graph(crimePPP_Drug)
```

### Calculating Distance (Nearest Neighborhood) of Crimes Patterns

```{r}
#Given two sets of points, this function computes 
#the Euclidean distance from each point in the first 
#set to each point in the second set, and returns a 
#matrix containing these distances. From this we take 
#the average of the distances in order to learn the 
#average distance between the types of crimes to each 
#other. If the value is higher than the rest this means 
#that these crimes occur farther away from each other.

cross_distance_table <- function(crime_type_dist){
  crime_Theft_Dist <- mean(crossdist(crime_type_dist[1:10000],crimePPP_Theft[1:10000]))
  #print(paste0("Theft Complete \n "))
  crime_Assualt_Dist <- mean(crossdist(crime_type_dist[1:10000],crimePPP_Assault[1:10000]))
  #print(paste0("Assualt Complete \n "))
  crime_Battery_Dist <- mean(crossdist(crime_type_dist[1:10000],crimePPP_Battery[1:10000]))
  #print(paste0("Battery Complete \n "))
  crime_Drug_Dist <- mean(crossdist(crime_type_dist[1:10000],crimePPP_Drug[1:10000]))
  #print(paste0("Drug Complete \n "))
  crime_Damage_Dist <- mean(crossdist(crime_type_dist[1:10000],crimePPP_Damage[1:10000]))
  #print(paste0("Damage Complete \n "))
  
  crime_distance_df <- data.frame(Assualt = as.numeric(0),Theft = as.numeric(0),Battery = as.numeric(0),Drug = as.numeric(0),Damage = as.numeric(0))
  
  crime_distance_df$Theft <- crime_Theft_Dist
  crime_distance_df$Assualt <- crime_Assualt_Dist
  crime_distance_df$Battery <- crime_Battery_Dist
  crime_distance_df$Drug <- crime_Drug_Dist
  crime_distance_df$Damage <- crime_Damage_Dist
  #print(paste0("Ready to make DataFrame  \n "))
  
  crime_distance_df
}


Theft_Dist_Comparison_DF <- cross_distance_table(crimePPP_Theft)
Assualt_Dist_Comparison_DF <- cross_distance_table(crimePPP_Assault)
Battery_Dist_Comparison_DF <- cross_distance_table(crimePPP_Battery)
Drug_Dist_Comparison_DF <- cross_distance_table(crimePPP_Drug)
Damage_Dist_Comparison_DF <- cross_distance_table(crimePPP_Damage)



Theft_Dist_Comparison_DF
Assualt_Dist_Comparison_DF
Battery_Dist_Comparison_DF
Drug_Dist_Comparison_DF
Damage_Dist_Comparison_DF


```

### Plotting G-Function to identify whether there is clustering in each crime pattern
 
```{r}

```

### Spatial Modeling of Crime Data


## Temporal Analysis

### Time Series Forecasting with Recurrent Neural Networks

[EXPLAIN WHY WE ARE UTILIZING RNN]

#### Long SHort-Term Memory Units (LSTMs)

## DBSCAN Analysis

[INSERT EXPLAINATION WHY WE ARE USING DBSCAN - DUE TO US FINDING THAT THERE ARE CLUSTERING WITHIN OUR CRIME PATTERNS WE NOW WANT TO INVESTIGATE THESE CLUSTERS AND VISUALIZE IF THEY ARE THE SAME CLUSTERS FOUND IN POINT PATTERN ANALYSIS]

[WHAT IS DBSCAN - What is it used for in spatial clustering ]

### Attempting 
```{r include=FALSE}
# lon <- crime_spatial_df$Longitude
# lat <- crime_spatial_df$Latitude
# xrange <- range(lon,na.rm = T)
# yrange <- range(lat,na.rm = T)
# crime_ppp <- ppp(lon,lat,xrange,yrange,marks = as.factor(crime_spatial_df$crime))
```

```{r}
# plot(crime_ppp, main = "Our Crime PPP Object")
```

```{r}
# qc <- quadratcount(crime_ppp)
# plot(qc)
```

